On the Synchronization Power of the Openflow Data Plane

### 2015-04-19 inband-sync.tex
small changes and comments

Given this memory interface, we can immediately implement 
a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
$\textit{old}$ and $\textit{new}$, reads the register and, if the
value of it is $\textit{old}$, replaces it with $\textit{new}$ and
returns $\top$; otherwise, $\bot$ is returned and
we say that the CAS operation \emph{fails} (which causes an abort of
the invoking transaction).

@@

In order to check if a value is written in a given address, 
we try to add it the same way as in the \memwrite operation,  
but we also set the \texttt{\checko} flag. If the value is not there,
our action differs from the action of the existing entry, 
thereby inflicting a failure.
In case the value is there, our flow entry replaces the existing flow entry 
with an identical one.
In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
of \emph{Compare-And-Set (CAS)}, using the two last operations.

@@

As a second example, let us consider a simple policy composition protocol that
allows a set of  controllers to concurrently update switch  \emph{policies}, i.e., sets of
effective flow tables rules, under the
condition that these rules can be composed with the currently installed
policy. 
Let us assume that the composition predicate is encoded in a user-specific
$\textit{apply\_update}()$ function~\cite{cpc}.

Algorithm~\ref{alg:simple-update} is based on the bundle feature and associates each policy with a \emph{policy identifier (pid)}.
In the algorithm, the controller first reads the currently installed
policy, then chooses a new unique policy identifier \emph{pid}, checks whether the new
policy and the current policy can be composed and, if so, replaces the
current policy with the composition. If, meanwhile, a new
policy (with a different identifier) has been installed, the update
fails and takes no effect.

We use increasing numbers as policies
IDs and use them in order to make sure that the new policy, with id
$x$, replaces the (what considered as) current policy, with id
$x-1$. Considering the policies IDs, we send the policy installation
commands as part of a bundle that also tries to perform
$CAS(x-1,x)$, thereby insuring that the commands are
performed iff the current policy is still there and preventing other
controllers to replace the new policy with another without observing it.

For brevity, we omit explicitly putting the $\textbf{bundle}\{...\}$ construct
around each of the transactional operation used in the algorithm.
Indeed, since \textit{claim} and \textit{unclaim} operations cannot
cause the corresponding transaction, we can assume that all
``singleton'' (consisting of one operation only)  transactions
containing such operations always commit.

@@

It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
 obtain a unique \emph{id}. FIXME: to be continued...
 \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.

@@

\textbf{TODO: fix the pseudocode, make the use of afdd/delete
  operations uniform, make all switch operations transactional}

\begin{algorithm}[t]
    \caption{Policy update with only CAS}
    \label{alg:simple-update}
    \begin{algorithmic}[1]
    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State $\pid,policy\gets$ \textbf{read} switch configuration
 			\State $cmds\gets U(policy)$
 			\startTxn
	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$
	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq \bot$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\begin{algorithm}[t]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
		 	\State $\pid,claims,policy\gets$ \textbf{read} switch configuration
 			\State $claim(\pid)$
 			\State $\pid 2,claims,policy\gets$ \textbf{read} switch configuration
 			\If {$\pid\neq \pid 2$}
	 			\State $unclaim(\pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $my\_id\gets$ choose a number from $C\setminus claims$
 			\State $cmds\gets U(policy)$
 			\startTxn
 				\State $\claimcheck(my\_id)$
	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$
	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(\pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\bot$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

>>>>

Given this memory interface, we can immediately implement 
a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
$\textit{old}$ and $\textit{new}$, reads the register and, if the
value of it is $\textit{old}$, replaces it with $\textit{new}$ and
returns $\top$; otherwise, $\bot$ is returned and
we say that the CAS operation \emph{fails} (which causes an abort of
the invoking transaction). \liron{CAS requieres to run in a bundle while the other operations disscussed here are not. Maybe we should remove CAS to the consensus subsection as it may confuses here}

@@

In order to check if a value is written in a given address, 
we try to add it the same way as in the \memwrite operation,  
but we also set the \texttt{\checko} flag. If the value is not there,
our action differs from the action of the existing entry, 
thereby inflicting a failure.
In case the value is there, our flow entry replaces the existing flow entry 
with an identical one.
In addition, we present in Algorithm \ref{alg:cas} a simple implementation 
of \emph{Compare-And-Set (CAS)}, using the two last operations. \liron{meybe move CAS to consensus subsection}

@@

As a second example, let us consider a simple policy composition protocol that
allows a set of  controllers to concurrently update switch  \emph{policies}, i.e., sets of
effective flow tables rules, under the
condition that these rules can be composed with the currently installed
policy. 
Let us assume that the composition predicate is encoded in a user-specific
$\textit{apply\_update}()$ function~\cite{cpc}.
\liron{TODO: fix the description of the composition predicate - receives a policy and return the commands. we also need to say how a normal controller code can be instrumented to use our paradigm}

Algorithm~\ref{alg:simple-update} is based on the bundle feature and associates each policy with a \emph{policy identifier (pid)}.
In the algorithm, the controller first reads the currently installed
policy and its id \emph{pid}, checks whether the new
policy and the current policy can be composed and, if so, replaces the
current policy with the composition. If, meanwhile, a new
policy (with a different identifier) has been installed, the update
fails and takes no effect. \liron{the term composition doesn't make sense now with the new user update function... }

We use increasing numbers as policies
IDs and use them in order to make sure that the new policy, with id
$pid+1$, replaces the (what considered as) current policy, with id
$pid$. Considering the policies IDs, we send the policy installation
commands as part of a bundle that also tries to perform
$CAS(pid,pid+1)$, thereby insuring that the commands are
performed iff the current policy is still there and preventing other
controllers to replace the new policy with another without observing it.

For brevity, we omit explicitly putting the $\textbf{bundle}\{...\}$ construct
around each of the transactional operation used in the algorithm.
Indeed, since \textit{claim} and \textit{unclaim} operations cannot
cause the corresponding transaction, we can assume that all
``singleton'' (consisting of one operation only)  transactions
containing such operations always commit. \liron{I think we can ommit this paragraph entirely - no need for bundle when it is not written}

@@

\begin{algorithm}[h]
    \caption{Policy update with only CAS}
    \label{alg:simple-update}
    \begin{algorithmic}[1]
    \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State $\pid,policy\gets$ \textbf{read} switch configuration
 			\State $cmds\gets U(policy)$
 			\startTxn
	 			\State $CAS(\pid_{\addr},\pid,\pid+1)$
	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq \bot$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

\hide{
It takes a set of rules $U$ (not yet a policy)and proceeds as follows: first, we seek to
 obtain a unique \emph{id}. FIXME: to be continued...
 \textbf{LS: I am not sure if I need to tell every step of the story or maybe it best to explain the main dif from the previous one similarly to what I just wrote next}.
}

@@

\textbf{TODO: fix the pseudocode, make the use of add/delete
  operations uniform, make all switch operations transactional}
  \liron{what does this TODO mean??}

@@

\begin{algorithm}[h]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
        \Require unsafe policy update function $U$, a predefined switch (memory) address for policy id: $\pid_{\addr}$, a predefined id space $C$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
		 	\State $\pid,claims,policy\gets$ \textbf{read} switch configuration
 			\State $claim(\pid)$
 			\State $\pid 2,claims,policy\gets$ \textbf{read} switch configuration
 			\If {$\pid\neq \pid 2$}
	 			\State $unclaim(\pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $my\_id\gets$ choose a number from $C\setminus claims$
 			\State $cmds\gets U(policy)$
 			\startTxn
 				\State $\claimcheck(my\_id)$
	 			\State $CAS(\pid_{\addr}, \pid,my\_id)$
	 			\State \textbf{send} $cmds$ %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(\pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\bot$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return
